---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setoptions, echo=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, warning=FALSE)
```

## Loading and preprocessing the data
```{r loading}
unzip ("activity.zip", exdir="./data")
df <- read.csv("./data/activity.csv")
unlink("./data", recursive=TRUE)
```
Now the data is loaded into variable df, let's preprocess the data
```{r preprocessing}
# Change date into date type
library(lubridate)
df$date <- ymd(df$date)
```

## What is mean total number of steps taken per day?
```{r mean_steps_per_day}
library(dplyr)
# Ignore missing values
df_clean <- df 
steps_per_day <- df %>% 
    filter(!is.na(steps)) %>%
    group_by(date) %>%
    summarise(sum=sum(steps))
```

The total number of steps taken per day is visualized through this historgram.
```{r visual}
library("ggplot2")
p <- ggplot(data=steps_per_day, aes(x=date, y=sum)) + geom_bar(stat="identity")
p + labs(x = "Date", y = "Total Steps", title=" Total Number of Steps Taken Per Day")
```


Report the mean and median total number of steps taken per day.
```{r report}
library(xtable)
xt <- xtable(data.frame(summary(data.frame(steps_per_day$sum))[3:4]))
print(xt, type="html")
```



## What is the average daily activity pattern?

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
